Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses

### 2023-01-20 RR_manuscript/RR_stage_2.tex
✨ - run RR diff

Based on the common practices currently in place within the field, we anticipated that researchers would use multilevel regression models, thus common measurements of effect size, such as Cohen's \(d\), might have been inappropriate.
Furthermore, \citet{aczel2021} suggest that directly asking analysts to report standardized effect sizes could bias the choice of analyses towards types that more straightforwardly return a standardized effect.
Since the variables used by the analysis teams might have substantially differed in their measurement scales (e.g, Hertz for frequency vs.~milliseconds for duration) which was indeed the case, we have standardized all reported effects by refitting each \textsc{reported model} with centered and scaled continuous variables (\emph{z}-scores, i.e.~the observed values subtracted from the mean divided by the standard deviation) and sum-coded factor variables.
Each \textsc{standardized model} was fitted as a Bayesian regression model with Stan \citep{stan2021}, RStan \citep{stan2020a}, and brms \citep{burkner2017} in R \citep{R-base}.
Model refitting also constituted a way of validating the reported analyses, a step recommended by \citet{aczel2021}.
Details about the refitting procedure can be found at \url{https://many-speech-analyses.github.io/many_analyses/scripts/r/04_refit_workflow}.
Relative to the registered protocol, we made minor changes to the refitting procedure, specifically file and variable naming conventions and the use of treatment contrasts instead of sum coding.
All models converged (\(\hat{R}\) was approximately 1).
Of the models with divergent transitions (n = 10), the number of divergences ranged from 1 to 143 (143 represents 3.58\% of total number of samples), which the authors deemed not to be problematic.

@@

We had initially planned to also include a measure of conservativeness of the model specification, as the number of random/group-level effects included and the number of post-hoc changes to the acoustic measurements the teams reported to have carried out.
When fitting the model, we realized that the measure of conservativeness is related to the standard error of the estimates (i.e.~more group-level effects = higher standard error).
Moreover, there was no team that declared to have made post-hoc changes to the analyses, thus we decided against including these two preregistered predictors in the model.

@@

All relevant data, code, and materials have been publicly archived on the Open Science Framework (\url{https://osf.io/3bmcp/}).
Archived data include the original data set distributed to all analysts, any edited versions of the data analyzed by individual teams, and the data we analyzed with our meta-analyses, which include the standardized effect sizes, the statistics describing variation in model structure among analysis teams, and the anonymized answers to our questionnaires of analysts.
Similarly, we archived both the analysis code used for each individual analysis and the code from our meta-analyses.
We also archived copies of our survey instruments from analysts and peer-reviewers.
Further documents concerning the collaborative editing of the Registered Report can be found at \url{https://drive.google.com/drive/folders/1-DOcj1qtEkvWfzu_FrsxkIGfPS0DyLXB?usp=sharing}.

We excluded from our synthesis any individual analysis submitted after peer review (Phase 3) or those unaccompanied by analysis files without which it was not possible to follow the research protocol.
We also excluded any individual analysis that does not produce an outcome that could be interpreted as an answer to our primary question.
We also did no include analyses for which we could not extract standardized effect sizes. For a list of exclusion criteria, see Section Descriptive statistics below.

@@

Forty-six teams submitted their analyses by the established deadline.
Only analyses from which it was possible to extract an effect size were included in the meta-analysis.
Of the analyses submitted by the 46 teams, the initiating authors identified analyses from 33 teams to be eligible to be included in the meta-analytic model.
Reasons for exclusion were: use of Generalized Additive Models (4 teams) which do not lend themselves easily to the meta-analytic methods employed in this study, use of machine learning techniques (3 teams), use of typicality as the outcome variable/response (3 teams), or use of other methods that returned statistics that could not be included in the meta-analytic model.
Note that due to the unforeseen variability across teams, the latter exclusion criteria were not preregistered and were applied after having seen all analytic strategies.

@@

The included analyses were provided by 33 teams, comprising 120 analysts, with a median of 3.0 individuals per team.
Upon sign-up, we collected background information from each analyst through the intake form, which was administered during Phase 1, prior to the data being released to the teams.
Analysts had a median of 5.4 years of experience after completing their PhD, ranging from -3.8 years, i.e.~PhD students (or less experienced) to 12.4 years, suggesting that, on average, analysts were experienced researchers.
The analysts' prior belief in the effect under investigation, on a scale from 0 to 100, ranged from 46.4 to 92.0 with a median of 70.0.
We take this to suggest that, overall, analysts had a rather high positive prior belief in the investigated relationship between acoustics and word combination typicality.

At the end of Phase 2 (primary data analysis), the teams had submitted a grand total of 115 individual models (including 192 critical model coefficients, given that some models returned more than one critical coefficient) to answer the research question, with a median of 3 models per team.
Table \ref{tab:descriptives-table} provides a summary of the contributing teams and their analyses.

@@

The analytic teams differed in their approach to the acoustic analysis of the speech signal, including choices related to specific acoustic measures, the temporal window used, and how the measures were transformed.
Thirty-seven percent of the models used f0 as the outcome variable, 33\% used a measure of duration, 13\% used vowel formants, 15\% intensity, and 3\% other measures.

Forty-five percent of models used acoustic measures taken at the level of the segment (e.g.~comparing the acoustic profile of a vowel), 45\% from the word level (e.g.~comparing the acoustic profile of \emph{Banane} `banana'), 3\% at the level of the phrase (e.g.~the noun phrase including determiner and adjective, e.g.~``the green banana''), 3\% from the whole sentence, and 3\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of 55 different measurement specifications.
For example, if we consider those analyses that target f0, we find that it is operationalized in many different ways including the minimum, the maximum, the mean, the median, as a range in an interval or a ratio between two intervals.
The measurement is sometimes taken from the interval of a vowel in the article, the adjective or noun; it is sometimes taken from the word interval of the article, adjective or noun; or it is taken from either the noun phrase interval or the entire sentence.
Some of these measures were normalized relative to other elements in the sentence or relative to the speaker.

@@

The mean of the number of different predictors included in teams' models was 2 (defined as variables or columns in the data table).
This means that, in addition to the critical predictor (typicality of the adjective noun combinations), models had on average one additional predictor (range = 1 - 5).
Possible information that was used as predictors included the information structure of the sentence, trial number, semantic dimensions of the referent, part of speech, and speaker gender.

The data given to the teams allowed them to operationalize the predictor of interest, word typicality, in different ways.
Among the possible operationalizations, 69\% of models contained typicality as a categorical variable (e.g.~atypical vs.~typical), 28\% used a continuous typicality scale from 0-100 by calculating the mean typicality for each word combination as obtained from the norming study, while 3\% of the models used the median typicality rating.
Note that the design of the experiment alongside its description indicated that the experiment was designed to categorically operationalize typicality. This possibly explains the analysts' strong preference.

The majority of models were run within a frequentist framework (84\%).
Sixteen percent were run within a Bayesian framework.
While teams almost exclusively used linear models to analyze their data (98\%), teams differed drastically in how they accounted for dependencies within the data.

The data contains several dependencies between data points, with multiple data points coming from the same subject and with multiple data points being associated with the same adjective or noun.
An appropriate way to account for this non-independence is by using models that include so-called random or group level effects \citep[e.g.,][]{gelman2006data, schielzeth2009conclusions}, variably known as mixed-effects, hierarchical, multi-level, or nested models (among other names).
Nine percent of the linear models specified no random effects at all (without pooling their data), effectively ignoring these non-independences \citep{hurlbert1984pseudoreplication}.
Sixty-two percent specified random intercepts only, and 29\% specified both random intercepts and random slopes to account for the non-independence.
On average, teams that specified random effects included 2.5 random terms in their models.
Based on statistical framework, type of model, distribution family, fixed terms, and not including random effects, there were a total of 52 different model specifications.

When considering both acoustic and statistical analyses, we have found a grand total of 119 different analytic pipelines.
In other words, each individual analysis submitted was unique.

@@

The question arises whether these unique analysis pipelines led to different conclusions.
Thirteen teams out of the thirty-three (39.4\%) reported to have found at least one statistically reliable effect (based on the inferential criteria they specified).
Of the 192 critical model coefficients, 45 were claimed to show a statistically reliable effect (23.4\%).

@@

\caption{\label{tab:descriptives-table}Descriptive statistics of teams, acoustic analyses, and statistical analyses included in the meta-analysis. The data set included analyses from 33 teams and  120 analysts.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llrr}
\toprule
\bf{Team characteristics} &    & \bf{Range} & \bf{Median}\\
\midrule
 & Team size & 1.0 -- 12.0 & 3.0\\
 & Years after PhD & -3.8 -- 12.4 & 5.4\\
 & Prior belief & 46.4 -- 92.0 & 70.0\\
 & Acoustic analysis peer rating & 41.2 -- 88.3 & 73.8\\
 & Statistical analysis peer rating & 33.0 -- 93.3 & 73.2\\
 & Overall peer rating & 39.0 -- 88.7 & 70.8\\
 &  &  \vphantom{5} & \\
\bf{Acoustic analyses} &  & \bf{n} & \bf{\%}\\
\midrule
\hspace{1em}Outcome & F0 & 44 & 37\\
 & Duration & 39 & 33\\
 & Intensity & 18 & 15\\
 & Formants & 15 & 13\\
 & Other & 3 & 3\\
 &  &  \vphantom{4} & \\
\hspace{1em}Temporal window & Segment & 54 & 46\\
 & Word & 53 & 45\\
 & Sentence & 4 & 3\\
 & Phrase & 3 & 3\\
 & Other & 4 & 3\\
 &  &  \vphantom{3} & \\
\hspace{1em}Typicality operationalization & Categorical & 82 & 69\\
 & Continuous (mean) & 33 & 28\\
 & Continuous (median) & 3 & 3\\
 &  &  \vphantom{2} & \\
\bf{Statistical analyses} &  & \bf{n} & \bf{\%}\\
\midrule
\hspace{1em}Framework & Frequentist & 100 & 84\\
 & Bayesian & 19 & 16\\
 &  &  \vphantom{1} & \\
\hspace{1em}Model & Linear model & 117 & 98\\
 & GAM & 1 & 1\\
 & Other & 1 & 1\\
 &  &  & \\
 &  & \bf{Range} & \bf{Median}\\
\hspace{1em}N & Models & 1 -- 16 & 3\\
 & Predictors & 1 -- 5 & 2\\
 & Random terms & 1 -- 10 & 2\\
\hspace{1em} & \hspace{1em}Intercept & \hspace{1em}1 -- 10 & \hspace{1em}2\\
\hspace{1em} & \hspace{1em}Slope & \hspace{1em}0 -- 4 & \hspace{1em}0\\
\bottomrule
\end{tabular}}
\end{table}

@@

Teams reviewed each others' acoustic and statistical analyses.
The mean rating of the acoustic analyses, on a scale from 0 to 100, is 71.5 (SD = 13.5).
The mean rating of the statistical analysis is 69.4 (SD = 15.9).
For reference, as mentioned in the Methods section, a score of 75 was defined as ``an imperfect analysis but the needed changes are unlikely to dramatically alter the final interpretation'', indicating that on average reviewers judged the provided analyses to be appropriate, although ``imperfect''.

@@

According to the preregistered meta-analytic model, the group-level standard deviation for teams is between 0.03 and 0.07 standard units at 95\% credibility.
In other words, the estimated range of variation across teams lies somewhere between \(\pm 0.06\) (0.03 * 1.96) and \(\pm 0.13\) (0.07 * 1.96) standard units with 95\% credibility.

\textbf{Non-preregistered}. However, in our preregistration we did not take into account that teams might submit multiple analyses/models which, if unaccounted for, violates the independence assumption.
Teams were explicitly instructed to only submit one effect size without enforcing it.
As a result, some teams followed the instruction and submitted only one model while others submitted multiple models.
To account for this added layer of dependency, we have run a model with team and model ID nested within team as group-level effects (\texttt{(1\textbar{}team)\ +\ (1\textbar{}team:model\_id)}), which allows us to estimate both the between-team variation and the between-analysis variation.
This analysis was not preregistered and should thus be interpreted with caution.\footnote{Note that before fitting this model, we fitted a separate one in which model ID was the only (non-nested) group-level effect.
  The estimated group-level effect of model ID is identical to that of the nested model, so we will not discuss it further.}

The nested model yields a posterior 95\% CrI for between-team variability of 0 to 0.04 standard units (\(\beta\) = 0.02, SD = 0.01), corresponding to a mean deviation range of about \(\pm 0\) to \(\pm 0.1\) standard units and 95\% probability.
The posterior 95\% CrI for between-analysis variability (nested within teams) is 0.11 to 0.14 standard units (\(\beta\) = 0.132, SD = 0.01).
For the sake of illustration, these would correspond to an estimate of between-model variability in segment and word durations that ranges between 7 to 14 ms for segments and between 7 and 33 ms for words at 95\% credibility.
We interpret these values in more details in the Discussion section.

@@

After assessing the variation between teams and analyses, we now turn to the meta-analytic estimate of the effect of typicality on the acoustic realization of sentences with adjective-noun combinations.
The meta-analytic model estimates the range of probable values of the standardized effect size to be between -0.026 and 0.016 standard units (95\% CrI, mean = -0.005).
In other words, our best guess is that speakers might not encode typicality in the acoustic signal (e.g.~by duration, f0, etc,) or, if they do, they do so by a maximum of \(\pm 0.03\) standard units.

\textbf{Non-preregistered}. As mentioned in the previous section, we have run an additional model, using team and model ID nested within team as group-level effects.
In this non-preregistered model, the meta-analytic intercept estimate is between -0.016 and 0.03 standard units (95\% CrI, \(\beta\) = 0.008).
This suggests that the acoustic measures of typical word combinations are 0.02 standard units lower to 0.03 standard units higher than the measures of atypical word combinations, at 95\% confidence.
This result is qualitatively similar to the results obtained in the preregistered model.

@@

Figure \ref{fig:plot-meta1} illustrates the individual intercepts for critical typicality coefficients across models and teams, sorted in ascending order based on their mean.
Given the nature and wide variety of acoustic operationalizations, there is no natural interpretation of the scale, so we cannot interpret the direction of estimates.
When looking at the raw estimates and their variance (grey triangles and lines), it is striking how much estimates differed.
Estimates ranged from -0.7 to 1.01 standard units.

While the majority of model estimates and their uncertainty after shrinkage (in black) yields inconclusive results (i.e.~are compatible with a point null hypothesis), there are 27 model estimates for which the 95\% credible interval does not contain zero (in orange, 14\%).

@@

We have run a model as described in Section Analytic and researcher-related predictors affecting effect sizes above.
Figure \ref{fig:plot-meta2}, panel C, displays the coefficients for all predictors alongside their 80\% and 95\% credible intervals.
The model suggests that most team-specific predictors yield very small deviations from the meta-analytic estimate and their 95\% credible intervals include zero, leaving us highly uncertain about their direction.
Neither analysts' prior beliefs in the phenomenon (\(\beta\) = -0.01, 95\% CrI = {[}-0.04, 0.01{]}), nor their seniority in terms of years after completing their PhD (\(\beta\) = 0.01, 95\% CrI = {[}-0.02, 0.04{]}) seem to affect model estimates.
Similarly, the evaluation of the quality of the analysis from their peers yielded a rather small effect magnitude, again characterized by large uncertainty (\(\beta\) = 0.02, 95\% CrI = {[}-0.01, 0.05{]}).
Interestingly, the model uniqueness, i.e.~how unique the choice and combination of predictors are, affects the analysts' estimate, with more unique models producing higher positive estimates (\(\beta\) = 0.04, 95\% CrI = {[}0.02, 0.07{]}).

@@

\begin{figure}
\includegraphics[width=1\linewidth]{../figs/alltogether} \caption{The effects of analytic and researcher-related predictors on the reported standardized effect sizes. (A) Posterior samples for the four most frequent outcome variables; (B) Posterior samples for the four most frequent temporal windows: Black points indicate medians; shaded areas represents 50/80/95\% highest density intervals. (C) Mean posterior samples (white circles) and 80/95\% credible intervals for all predictors grouped into predictors related to (1) temporal window, (2) outcome variable, and (3) team/analysis.}\label{fig:plot-meta2}
\end{figure}

@@

We gave 46 analyst teams the same speech data set to answer the same research question: \emph{Do
speakers acoustically modify utterances to signal atypical word combinations?}
In order to answer this question, teams had to interpret the research question by
operationalizing constructs within multidimensional signals, operationalizing and choosing appropriate model predictors, and constructing appropriate statistical models.
This complex process has led to a vast ``garden of forking paths'', i.e.~to a wide range of combinations of possible analytic decisions.
The submitted analyses exhibited at least 52 unique ways of operationalizing the acoustic signal alongside 55 unique ways of constructing the statistical model.
By multiplying the numbers of acoustic and model specifications, there are in principle 2860 possible unique combinations.
Note that this is a conservative estimate of the number of possible analytic choices for our research question, ignoring many other degrees of freedom like e.g.~acoustic parameter extraction, outlier treatment, and transformations, all of which might have an impact on the final results \citep{breznau2021observing}.

Different analysis paths led to different categorical conclusions with 39.4\% of teams reported to have found at least one statistically reliable effect.
To gain a better understanding of whether the observed quantitative variability can result in theoretically different claims, we will contextualize them in actual acoustic measures.
We calculated the standard deviation of a selection of acoustic measurements, as submitted by the analysis teams: duration, f0 and intensity, taken from different time windows.
These standard deviations can be considered as a coarse indication of the variability in the obtained acoustic measures.
We can now use these values to interpret the meta-analytic estimates, which are in standardized units, by transforming the standardized units to measures of duration, f0 and intensity.\footnote{Note that these categories necessarily refer to a variegated set of measures, for example the domain ``word'' includes words that differed along several dimensions, including their length and their metrical structure.}

>>>>

Based on the common practices currently in place within the field, we anticipated that researchers would use multilevel regression models, thus common measurements of effect size, such as Cohen's \(d\), might have been inappropriate.
Furthermore, \citet{aczel2021} suggest that directly asking analysts to report standardized effect sizes could bias the choice of analyses towards types that more straightforwardly return a standardized effect.
Since the variables used by the analysis teams might have substantially differed in their measurement scales (e.g, Hertz for frequency vs.~milliseconds for duration) which was indeed the case, we have standardized all reported effects by refitting each \textsc{reported model} with centered and scaled continuous variables (\emph{z}-scores, i.e.~the observed values subtracted from the mean divided by the standard deviation) and sum-coded factor variables.
Each \textsc{standardized model} was fitted as a Bayesian regression model with Stan \citep{stan2021}, RStan \citep{stan2020a}, and brms \citep{burkner2017} in R \citep{R-base}.
Model refitting also constituted a way of validating the reported analyses, a step recommended by \citet{aczel2021}.
Details about the refitting procedure can be found at \url{https://many-speech-analyses.github.io/many_analyses/scripts/r/04_refit_workflow}.

@@

We had initially planned to also include a measure of conservativeness of the model specification, as the number of random/group-level effects included and the number of post-hoc changes to the acoustic measurements the teams reported to have carried out.
When fitting the model, we realized that the measure of conservativeness is related to the standard error of the estimates (i.e.~more group-level effects = higher standard error).
Moreover, there was no team that declared to have made post-hoc changes to the analyses, this we decided against including these two preregistered predictors in the model.

@@

All relevant data, code, and materials have been publicly archived on the Open Science Framework (\url{https://osf.io/3bmcp/}).
Archived data include the original data set distributed to all analysts, any edited versions of the data analyzed by individual teams, and the data we analyzed with our meta-analyses, which include the standardized effect sizes, the statistics describing variation in model structure among analysis teams, and the anonymized answers to our questionnaires of analysts.
Similarly, we archived both the analysis code used for each individual analysis and the code from our meta-analyses.
We also archived copies of our survey instruments from analysts and peer-reviewers.

We excluded from our synthesis any individual analysis submitted after peer review (Phase 3) or those unaccompanied by analysis files without which it was not possible to follow the research protocol.
We also excluded any individual analysis that does not produce an outcome that could be interpreted as an answer to our primary question.
For a list of exlcusion criteria, see Section Descriptive statistics below.

@@

Forty-six teams submitted their analyses by the established deadline.
Only analyses from which it was possible to extract an effect size were included in the meta-analysis.
Of the analysis submitted by the 46 teams, the initiating authors identified analyses from 30 teams to be eligible to be included in the meta-analytic model.
Reasons for exclusion were: use of Generalized Additive Models (4 teams) which do not lend themselves easily to the meta-analytic methods employed in this study, use of machine learning techniques (3 teams), use of typicality as the outcome variable/response (3 teams), or use of other methods that returned statistics that could not be included in the meta-analytic model.

@@

The included analyses were provided by 30 teams, comprising 109 analysts, with a median of 3.0 individuals per team.
Upon sign-up, we collected background information from each analyst through the intake form, which was administered during Phase 1, prior to the data being released to the teams.
Analysts had a median of 5.8 years of experience after completing their PhD, ranging from -3.8 years, i.e.~PhD students (or less experienced) to 12.4 years, suggesting that, on average, analysts were experienced researchers.
The analysts' prior belief in the effect under investigation, on a scale from 0 to 100, ranged from 48.5 to 92.0 with a median of 70.0.
We take this to suggest that, overall, analysts had a rather high positive prior belief in the investigated relationship between acoustics and word combination typicality.

At the end of Phase 2 (primary data analysis), the teams had submitted a grand total of 104 individual models (including 170 critical model coefficients, given that some models returned more than one critical coefficient) to answer the research question, with a median of 3 models per team.
Table \ref{tab:descriptives-table} provides a summary of the contributing teams and their analyses.

@@

The analytic teams differed in their approach to the acoustic analysis of the speech signal, including choices related to specific acoustic measures, the temporal window used, and how the measures were transformed.
Thirty-six percent of the models used f0 as the outcome variable, 35\% used a measure of duration, 12\% used vowel formants, 14\% intensity, and 3\% other measures.

Forty-five percent of models used acoustic measures taken at the level of the segment (e.g.~comparing the acoustic profile of a vowel), 43\% from the word level (e.g.~comparing the acoustic profile of \emph{Banane} `banana'), 3\% at the level of the phrase (e.g.~the noun phrase including determiner and adjective, e.g.~``the green banana''), 4\% from the whole sentence, and 4\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of 52 different measurement specifications.
For example, if we consider those analyses that target f0, we find that it is operationalized in many different ways including the minimum, the maximum, the mean, the median, as a range in an interval or a ratio between two intervals.
The measurement is sometimes taken from the interval of a vowel in the article, the adjective or noun; it is sometimes taken from the word interval of the article, adjective or noun; or it is taken from either the noun phrase interval or the entire sentence.
Some of these measures were normalized relative to other elements in the sentence or relative to the speaker.

@@

The mean of the number of different predictors included in teams' models was 2 (defined as variables or columns in the data table).
This means that, in addition to the critical predictor (typicality of the adjective noun combinations), models had on average one additional predictor (range = 1-\texttt{r}round(predictor\_n\$max\_predictor,1)`).
Possible information that was used as predictors included the information structure of the sentence, trial number, semantic dimensions of the referent, part of speech, and speaker gender.

The data given to the teams allowed them to operationalize the predictor of interest, word typicality, in different ways.
Among the possible operationalizations, 72\% of models contained typicality as a categorical variable (e.g.~atypical vs.~typical), 24\% used a continuous typicality scale from 0-100 by calculating the mean typicality for each word combination as obtained from the norming study, while 3\% of the models used the median typicality rating.
Note that the design of the experiment alongside its description indicated that the experiment was designed to categorically operationalize typicality. This possibly explains analysts' strong preference.

The majority of models were run within a frequentist framework (83\%).
Seventeen percent were run within a Bayesian framework.
While teams almost exclusively used linear models to analyze their data (98\%), teams differed drastically in how they accounted for dependencies within the data.

The data contains several dependencies between data points, with multiple data points coming from the same subject and with multiple data points being associated with the same adjective or noun.
An appropriate way to account for this non-independence is by using models that include so-called random or group level effects \citep[e.g.,][]{gelman2006data, schielzeth2009conclusions}, variably known as mixed-effects, hierarchical, multi-level, or nested models (among other names).
Eight percent of the linear models specified no random effects at all (without pooling their data), effectively ignoring these non-independences \citep{hurlbert1984pseudoreplication}.
Sixty-five percent specified random intercepts only, and 27\% specified both random intercepts and random slopes to account for the non-independence.
On average, teams that specified random effects included 2.5 random terms in their models.
Based on statistical framework, type of model, distribution family, fixed terms, and not including random effects, there were a total of 47 different model specifications.

When considering both acoustic and statistical analyses, we have found a grand total of 104 different analytic pipelines.
In other words, each individual analysis submitted was unique.

@@

The question arises whether these unique analysis pipelines led to different conclusions.
Nine teams out of the thirty (30\%) reported to have found at least one statistically reliable effect (based on the inferential criteria they specified).
Of the 104 submitted models, 37 were reported to show a statistically reliable effect (21.8\%).

@@

\caption{\label{tab:descriptives-table}Descriptive statistics of teams, acoustic analyses, and statistical analyses included in the meta-analysis. The data set included analyses from 30 teams and  109 analysts.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llrr}
\toprule
\bf{Team characteristics} &    & \bf{Range} & \bf{Median}\\
\midrule
 & Team size & 1.0 -- 12.0 & 3.0\\
 & Years after PhD & -3.8 -- 12.4 & 5.8\\
 & Prior belief & 48.5 -- 92.0 & 70.0\\
 & Acoustic analysis peer rating & 41.2 -- 88.3 & 75.2\\
 & Statistical analysis peer rating & 33.0 -- 93.3 & 74.1\\
 & Overall peer rating & 39.0 -- 88.7 & 72.2\\
 &  &  \vphantom{5} & \\
\bf{Acoustic analyses} &  & \bf{n} & \bf{\%}\\
\midrule
\hspace{1em}Outcome & F0 & 37 & 36\\
 & Duration & 36 & 35\\
 & Intensity & 15 & 14\\
 & Formants & 13 & 12\\
 & Other & 3 & 3\\
 &  &  \vphantom{4} & \\
\hspace{1em}Temporal window & Segment & 47 & 46\\
 & Word & 45 & 44\\
 & Sentence & 4 & 4\\
 & Phrase & 3 & 3\\
 & Other & 4 & 4\\
 &  &  \vphantom{3} & \\
\hspace{1em}Typicality operationalization & Categorical & 75 & 73\\
 & Continuous (mean) & 25 & 24\\
 & Continuous (median) & 3 & 3\\
 &  &  \vphantom{2} & \\
\bf{Statistical analyses} &  & \bf{n} & \bf{\%}\\
\midrule
\hspace{1em}Framework & Frequentist & 86 & 83\\
 & Bayesian & 18 & 17\\
 &  &  \vphantom{1} & \\
\hspace{1em}Model & Linear model & 102 & 98\\
 & GAM & 1 & 1\\
 & Other & 1 & 1\\
 &  &  & \\
 &  & \bf{Range} & \bf{Median}\\
\hspace{1em}N & Models & 1 -- 16 & 3\\
 & Predictors & 1 -- 5 & 2\\
 & Random terms & 1 -- 10 & 2\\
\hspace{1em} & \hspace{1em}Intercept & \hspace{1em}1 -- 10 & \hspace{1em}2\\
\hspace{1em} & \hspace{1em}Slope & \hspace{1em}0 -- 4 & \hspace{1em}0\\
\bottomrule
\end{tabular}}
\end{table}

@@

Teams reviewed each others' acoustic and statistical analyses.
The mean rating of the acoustic analyses, on a scale from 0 to 100, is 71.5 (SD = 13.9).
The mean rating of the statistical analysis is 69.5 (SD = 16.5).
For reference, as mentioned in the Methods section, a score of 75 was defined as ``an imperfect analysis but the needed changes are unlikely to dramatically alter the final interpretation'', indicating that on average reviewers judged the provided analyses to be appropriate, although ``imperfect''.

@@

According to the preregistered meta-analytic model, the group-level standard deviation for teams is between 0.03 and 0.07 standard units at 95\% credibility.
In other words, the estimated range of variation across teams lies somewhere between \(\pm 0.06\) (0.03 * 1.96) and \(\pm 0.14\) (0.07 * 1.96) with 95\% credibility.''

\textbf{Non-preregistered}. However, in our preregistration we did not take into account that teams might submit multiple analyses/models which, if unaccounted for, violates the independence assumption.
Teams were explicitly instructed to only submit one effect size without enforcing it.
As a result, some teams followed the instruction and submitted only one model while others submitted multiple models.
To account for this added layer of dependency, we have run a model with team and model ID nested within team as group-level effects (\texttt{(1\textbar{}team)\ +\ (1\textbar{}team:model\_id)}), which allows us to estimate both the between-team variation and the between-analysis variation.
This analysis was not preregistered and should thus be interpreted with caution.\footnote{Note that before fitting this model, we fit a separate one in which model ID was the only (non-nested) group-level effect.
  The estimated group-level effect of model ID is identical to that of the nested model, so we will not discuss it further.}

The nested model yields a posterior 95\% CrI for between-team variability of 0 to 0.05 standard units (\(\beta\) = 0.02, SD = 0.01), corresponding to a mean deviation range of about \(\pm 0\) to \(\pm 0.1\) standard units and 95\% probability.
The posterior 95\% CrI for between-analysis variability (nested within teams) is 0.11 to 0.15 standard units (\(\beta\) = 0.132, SD = 0.01).
For the sake of illustration, these would correspond to an estimate of between-model variability in milliseconds (if looking for example at segment duration) that ranges between 7.4 and 15.1 ms at 95\% credibility.
We interpret these values in more details in the Discussion section.

@@

After assessing the variation between teams and analyses, we now turn to the meta-analytic estimate of the effect of typicality on the acoustic realization of sentences with adjective-noun combinations.
The meta-analytic model estimates the range of probable values of the standardized effect size to be between -0.024 and 0.021 standard units (95\% CrI, mean = -0.001).
In other words, our best guess is that speakers might not encode typicality in the acoustic signal (e.g.~by duration, f0, etc,) or, if they do, they do so by a maximum of \(\pm 0.02\) standard units.

\textbf{Non-preregistered}. As mentioned in the previous section, we have run an additional model, using team and model ID nested within team as group-level effects.
In this non-preregistered model, the meta-analytic intercept estimate is between -0.013 and 0.04 standard units (95\% CrI, \(\beta\) = 0.014).
This suggests that the acoustic measures of typical word combinations are 0.01 standard units lower to 0.04 standard units higher than the measures of atypical word combinations, at 95\% confidence.
Relative to the preregistered model, this model suggests a somewhat positive effect of typicality (although small negative effects are also possible).

@@

Figure \ref{fig:plot-meta1} illustrates the individual intercepts for critical typicality coefficients across models and teams, sorted in ascending order based on their mean.
Given the nature and wide variety of acoustic operationalizations, there is no natural interpretation of the scale, so we cannot interpret the direction of estimates.
When looking at the raw estimates and their variance (grey triangles and lines), it is striking how much estimates differed.
Estimates ranged from -0.65 to 0.99 standard units.

While the majority of model estimates and their uncertainty after shrinkage (in black) yields inconclusive results (i.e.~are compatible with a point null hypothesis), there are 27 model estimates for which the 95\% credible interval does not contain zero (in blue, 16\%).

@@

We have run a model as described in Section Analytic and researcher-related predictors affecting effect sizes above.
Figure \ref{fig:plot-meta2}, panel C, displays the coefficients for all predictors alongside their 80\% and 95\% credible intervals.
The model suggests that most team-specific predictors yield very small deviations from the meta-analytic estimate and their 95\% credible intervals include zero, leaving us highly uncertain about their direction.
Neither analysts' prior beliefs in the phenomenon (\(\beta\) = 0.00, 95\% CrI = {[}-0.03, 0.03{]}), nor their seniority in terms of years after completing their PhD (\(\beta\) = 0.01, 95\% CrI = {[}-0.03, 0.04{]}) seem to affect model estimates.
Similarly, the evaluation of the quality of the analysis from their peers yielded a rather small effect magnitude, again characterized by large uncertainty (\(\beta\) = 0.01, 95\% CrI = {[}-0.02, 0.05{]}).
Interestingly, the model uniqueness, i.e.~how unique the choice and combination of predictors are, affects the analysts' estimate, with more unique models producing higher positive estimates (\(\beta\) = 0.05, 95\% CrI = {[}0.02, 0.07{]}).

@@

\begin{figure}
\includegraphics[width=1\linewidth]{../figs/alltogether} \caption{The effects of analytic and researcher-related predictors on the reported standardized effect sizes. (A) Posterior samples for the four most frequent outcome variables; (B) Posterior samples for the four most frequent temporal windows: Black points indicate medians; shaded areas represents 50/80/95\% highest density intervals. (C) Mean posterior samples (white circles) and 80/95\% credible intervals for all predictors.}\label{fig:plot-meta2}
\end{figure}

@@

We gave 46 analyst teams the same speech data set to answer the same research question: \emph{Do
speakers acoustically modify utterances to signal atypical word combinations?}
In order to answer this question, teams had to interpret the research question by
operationalizing constructs within multidimensional signals, operationalizing and choosing appropriate model predictors, and constructing appropriate statistical models.
This complex process has led to a vast ``garden of forking paths'', i.e.~to a wide range of combinations of possible analytic decisions.
The submitted analyses exhibited at least 47 unique ways of operationalizing the acoustic signal alongside 52 unique ways of constructing the statistical model.
By multiplying the numbers of acoustic and model specifications, there are in principle 2444 possible unique combinations.
Note that this is a conservative estimate of the number of possible analytic choices for our research question, ignoring many other degrees of freedom like e.g.~acoustic parameter extraction, outlier treatment, and transformations, all of which might have an impact on the final results \citep{breznau2021observing}.

Different analysis paths led to different categorical conclusions with 30\% of teams reported to have found at least one statistically reliable effect.
To gain a better understanding of whether the observed quantitative variability can result in theoretically different claims, we will contextualize them in actual acoustic measures.
We calculated the standard deviation of a selection of acoustic measurements, as submitted by the analysis teams: duration, f0 and intensity, taken from different time windows.
These standard deviations can be considered as a coarse indication of the variability in the obtained acoustic measures.
We can now use these values to interpret the meta-analytic estimates, which are in standardized units, by transforming the standardized units to measures of duration, f0 and intensity.\footnote{Note that these categories necessarily refer to a variegated set of measures, for example the domain ``word'' includes words that differed along several dimensions, including their length and their metrical structure.}

For example, for those analyses that investigated the duration of vowels (e.g.~the duration of the stressed vowel in \emph{Banane}), the reported duration measures exhibit standard deviations that range from 33.4 to 51.4 ms.
These standard deviations allow us to convert the meta analytic estimates into milliseconds by multiplying those values with the standard units.
The reported effect estimates from teams varied between -0.65 and 0.99 standard units, which corresponds to duration values ranging from -21.86 to 50.81 ms.
A more conservative approach is to convert the meta-analytic estimates of between-model variation, thus obtaining an estimate of between-model variability in milliseconds that ranges between 7.4 and 15.1 ms at 95\% credibility.\footnote{The calculation is thus: the minimum standard deviation of duration multiplied by the lower limit of the 95\% CrI of the between-models variability estimate, times 1.96 to obtain a 95\% CrI: \(33.4 * 0.11 = 3.8\), \(3.8 * 1.96 = 7.4\) ms; the maximum standard deviation of duration multiplied by the upper limit of the 95\% CrI of the between-models variability estimate, times 1.96: \(51.4 * 0.15 = 7.7\), \(7.7 * 1.96 = 15.1\) ms.}
